Confirmation of \srcfull{} as a long period radio transient

### 2025-06-04 paper/main.tex
Synched paper with Overleaf version

Pulses from \src{} were detected with the Murchison Widefield Array \citep[MWA;][]{Tingay2013}, the Australian SKA Pathfinder \citep[ASKAP;][]{2021PASA...38....9H}, MeerKAT \citep{2016mks..confE...1J}, and the Australia Telescope Compact Array (ATCA, \todo{CITATION??}), all radio interferometers capable of imaging.
A summary of all the observations containing pulse detections included in this work is summarised in \Tab~\ref{tbl:obs}.
In this section, we describe how each telescope was used to generate dynamic spectra of \src{}.

\begin{table*}
  \centering
  \caption{Observations summary}
  \label{tbl:obs}
  \begin{tabular}{lllcccc}
    \hline
    Date range & Telescope & Project & $\nu$ & $\Delta\nu$ & Obs. (num. $\times$) & Number of \\ 
    (MJD) & & & (MHz) & (MHz) & length (min) & pulses \\
    \hline 
    59965 & ASKAP & SB47253 (original discovery) & 887.5 & 288 & 12.1 & 1 \\
    59966 & MWA & G0060 & 161.92 & 15.36 & 6.7 & 1 \\
    %60084 & MWA & G0057 & 154.24 & 30.72 & N/A & $\le$ 1 \\
    60040  & ASKAP & SB49153 & 887.5 & 288 & 12.1 & 1 \\
    60092 - 60093 & MeerKAT & DDT-20230525-DD-01 & 1284 & 856 & 307 & 4 \\
    60463 - 60563 & MWA & G0080 & 200.32 & 30.72 & $16 \times 4.9$ & 16 \\
    60503 & ASKAP & SB63600 & 887.5 & 288 & 12.3 & 1 \\
    60572 & MWA & D0042 & 184.96 & 30.72 & $4 \times 4.9$ & 4 \\
    60587 & ATCA & \todo{?????} & 2100 & 2048 & 93.2 & 2 \\
    60592 - 60602 & MWA & D0042 & 200.32 & 30.72 & $17 \times 4.9$ & 17 \\
    60602 & MeerKAT & DDT-20241015-NH-01 & 812.8 & 495.3 & 10 & 1 \\
    \hline
  \end{tabular}
\end{table*}

@@

\todo{Akash to add some paragraphs here with details of processing ATCA data:}
\begin{itemize}
    \item Describe both Oct 4 and Oct 12 observations
    \item Mention that the Oct 12 observations did not yield any detections (even though there were four chances for it to do so!).
\end{itemize}

@@

To account for the asymmetry, we fit exponentially modified Gaussians to each pulse with the form,
\begin{equation}
  \emg(t) = A \frac{\sigma}{\tau}\sqrt{\frac{\pi}{2}}
    \exp\left(-\frac12 \left( \frac{t - \mu}{\sigma} \right)^2 \right)  \erfcx \left(\frac{1}{\sqrt{2}} \left(\frac{\sigma}{\tau} - \frac{t - \mu}{\sigma} \right) \right),
  \label{eqn:emg}
\end{equation}
where $\erfcx(X) = \exp(X^2) \erfc(X)$ is the scaled complimentary error function; $A$, $\mu$, and $\sigma$ are the height, location, and scale of an unscattered Gaussian pulse; and the scattering time scale was fixed at
\begin{equation}
      \tau(\nu) = 0.07\, \left(\frac{\nu}{\rm GHz}\right)^{-4} \, {\rm s},
\end{equation}
according to each observation's centre frequency.
A pulse stack containing all observed pulses and the corresponding unscattered pulses derived from the fits is shown in \Fig~\ref{fig:pulsestack}.

@@

No significant linear or circular polarised emission was detected in the 2023 MeerKAT pulses.
This is likely due to low S/N coupled with the source not being centred in the beam.
Much more careful imaging analysis would be needed to obtain reliable measurements or limits on the polarised components of those pulses.
We also find no significant polarised emission from the ATCA pulses. \todo{[Akash: Can you please verify this?]}

@@

In contrast, the ATCA observation taken on 2025-10-12 did not yield any detections (out of four possible predicted ToAs that occurred during that observation).
This is despite the fact that pulses were detected with ATCA just eight days prior with a similar noise level, as well as the fact that pulses were detected at the MWA within 24 hours of the ATCA observation (none of the MWA observations were coincident with the ATCA observation).
However, there is enough pulse-to-pulse variability to allow for the possibility that the pulses during the 2025-10-12 were present, but fell below the detection threshold of ATCA.

@@

The raw visibilities for the MWA, ASKAP, and MeerKAT data sets are available via the respective telescopes' data portals. 
The post-imaged dynamic spectra, analyses, and timing analysis data products (ToAs, pulse properties) are publicly available on GitHub\footnote{TODO: Make this available upon publication.}.

>>>>

Pulses from \src{} were detected with the Murchison Widefield Array \citep[MWA;][]{Tingay2013}, the Australian SKA Pathfinder \citep[ASKAP;][]{2021PASA...38....9H}, MeerKAT \citep{2016mks..confE...1J}, and the Australia Telescope Compact Array \citep[ATCA;][]{cabb}, all radio interferometers capable of imaging.
A summary of all the observations containing pulse detections included in this work is summarised in \Tab~\ref{tbl:obs}.
In this section, we describe how each telescope was used to generate dynamic spectra of \src{}.

\begin{table*}
  \centering
  \caption{Observations summary}
  \label{tbl:obs}
  \begin{tabular}{lllcccc}
    \hline
    Date range & Telescope & Project & $\nu$ & $\Delta\nu$ & Obs. (num. $\times$) & Number of \\ 
    (MJD) & & & (MHz) & (MHz) & length (min) & pulses \\
    \hline 
    59965 & ASKAP & SB47253 (original discovery) & 887.5 & 288 & 12.1 & 1 \\
    59966 & MWA & G0060 & 161.92 & 15.36 & 6.7 & 1 \\
    %60084 & MWA & G0057 & 154.24 & 30.72 & N/A & $\le$ 1 \\
    60040  & ASKAP & SB49153 & 887.5 & 288 & 12.1 & 1 \\
    60092 - 60093 & MeerKAT & DDT-20230525-DD-01 & 1284 & 856 & 307 & 4 \\
    60463 - 60563 & MWA & G0080 & 200.32 & 30.72 & $16 \times 4.9$ & 16 \\
    60503 & ASKAP & SB63600 & 887.5 & 288 & 12.3 & 1 \\
    60572 & MWA & D0042 & 184.96 & 30.72 & $4 \times 4.9$ & 4 \\
    60587 & ATCA & C3363 & 2100 & 2048 & 93.2 & 2 \\
    60592 - 60602 & MWA & D0042 & 200.32 & 30.72 & $17 \times 4.9$ & 17 \\
    60602 & MeerKAT & DDT-20241015-NH-01 & 812.8 & 495.3 & 10 & 1 \\
    \hline
  \end{tabular}
\end{table*}

@@

Following its re-detection in the ASKAP data, follow-up observations were conducted at L/S band with the ATCA telescope under the project code C3363 (P.I: Tara Murphy). \src\ was observed once on October 4, 2024, for a total of 1.5\,hrs, and again on October 12, 2024, for a total of 2.4\,hrs. At the beginning of each observation, 1934$-$638 was used to calibrate the bandpass, and during the observations, 1748$-$253 was used as a phase calibrator. Data were flagged, calibrated, and imaged using standard recipes from the Common Astronomy Software Applications package \citep[CASA;][]{casa}. The resulting sky model was subtracted from the visibilities (excluding the source of interest). These model-subtracted visibilities were phase-centered on the \src{} and averaged over the baselines to produce the dynamic spectra.

We identified two faint pulses (4--5 $\sigma$; see Table~\ref{tbl:toas}) in the observation taken on October 4. However, observations taken on October 12 did not yield any pulses (at 3-$\sigma$ level). We inspected the data around the expected time of arrival (using Equation~\ref{eqn:timing}) but did not find any pulses. The instantaneous noise level around the expected time of arrivals is close to 1\,mJy. If pulse-to-pulse flux density variations, similar to ones observed in MWA data (Table~\ref{tbl:toas}), also occur at L/S band, then even a factor 1.5--2 can be sufficient to explain the detection of pulse on October 4 and the absence of pulses on October 12, given that the source is faint at higher frequencies.

@@

To account for the asymmetry, we fit exponentially modified Gaussians to each pulse with the form,
\begin{equation}
  \emg(t) = A \frac{\sigma}{\tau}\sqrt{\frac{\pi}{2}}
    \exp\left(-\frac12 \left( \frac{t - \mu}{\sigma} \right)^2 \right)  \erfcx \left(\frac{1}{\sqrt{2}} \left(\frac{\sigma}{\tau} - \frac{t - \mu}{\sigma} \right) \right),
  \label{eqn:emg}
\end{equation}
where $\erfcx(X) = \exp(X^2) \erfc(X)$ is the scaled complementary error function; $A$, $\mu$, and $\sigma$ are the height, location, and scale of an unscattered Gaussian pulse; and the scattering time scale was fixed at
\begin{equation}
      \tau(\nu) = 0.07\, \left(\frac{\nu}{\rm GHz}\right)^{-4} \, {\rm s},
\end{equation}
according to each observation's centre frequency.
A pulse stack containing all observed pulses and the corresponding unscattered pulses derived from the fits is shown in \Fig~\ref{fig:pulsestack}.

@@

No significant linear or circular polarised emission was detected in the 2023 MeerKAT pulses.
This is likely due to low S/N coupled with the source not being centred in the beam.
Much more careful imaging analysis would be needed to obtain reliable measurements or limits on the polarised components of those pulses.
We also find no significant polarised emission from the ATCA pulses.

@@

In contrast, the ATCA observation taken on 2025-10-12 did not yield any detections (out of four possible predicted ToAs that occurred during that observation).
This is despite the fact that pulses were detected with ATCA just eight days prior with a similar instantaneous noise level (${\sim}1\,$mJy), as well as the fact that pulses were detected at the MWA within 24 hours of the ATCA observation (none of the MWA observations were coincident with the ATCA observation).
However, even a modest amount of pulse-to-pulse variability would allow for the possibility that the pulses during the 2025-10-12 were present, but fell below the detection threshold of ATCA.

@@

The raw visibilities for the MWA, ASKAP, and MeerKAT data sets are available via the respective telescopes' data portals.
The ATCA data is available on reasonable request to the authors, and will become publicly available from the relevant data portal by November, 2025.
The post-imaged dynamic spectra, analyses, and timing analysis data products (ToAs, pulse properties) are publicly available on GitHub\footnote{TODO: Make this available upon publication.}.
